---
title: "Analysis of Gapminder"
author: "YuanChen Ling"
date: "2026-09-23"
---
##Analyzing the Gapminder Dataset

In this post, I explore the Gapminder dataset to examine patterns in population, GDP per capita, and life expectancy across countries.

```{r}
library(tidyverse)
library(gapminder)
#| message: false
#| warning: false
gapminder
```

## Average Life Expectancy by Continent
To compare life expectancy across continents, I calculate the mean life expectancy for each continent.
```{r}
life_exp_by_continent <- gapminder |>
  group_by(continent) |>
  summarise(
    mean_life_exp = mean(lifeExp)
  ) |>
  arrange(desc(mean_life_exp))

life_exp_by_continent
```
## GDP per Capita and Life Expectancy

Next, I visualize the relationship between GDP per capita and life expectancy. Each point represents a country in a particular year, and colors distinguish continents.

```{r}
#| label: fig-gdp-lifeexp
#| fig-cap: "Relationship between GDP per capita and life expectancy across continents."

ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point(alpha = 0.6) +
  scale_x_log10() +
  labs(
    x = "GDP per Capita",
    y = "Life Expectancy",
    color = "Continent"
  ) +
  theme_minimal()
```
### Data Source

The data used in this analysis come from the `gapminder` R package. The dataset contains country-level information on life expectancy, population, and GDP per capita across multiple years.